---
title: "R Notebook"
output: html_notebook
---
 
# 0. Import Relevant Libraries

```{r}

library(dplyr)
library(e1071)
library(ggplot2)
library(gmodels)
library(lubridate)
library(NLP)
library(reshape2)
library(RColorBrewer)
library(scales)
library(SnowballC)
library(stringr)
library(syuzhet)
library(tm)
library(wordcloud)

```

# 1. Read in data sets

```{r}

#Read in December 2022 data
chatGPTDataDec <- read.csv("chatGPTNovDec2022.csv", stringsAsFactors = FALSE)

#Read in 2023 data
chatGPTData2023 <- read.csv("chatGPTJanMar2023.csv", stringsAsFactors = FALSE)

```

# 2. Initial data cleanup 

## a. December 2023 data

```{r}

#Convert created_at to time
chatGPTDataDec$created_at <- as.POSIXct(chatGPTDataDec$created_at, format = "%Y-%m-%d %H:%M:%S")

#Remove 11/30/22 to have only the month of December
chatGPTDataDec <- subset(chatGPTDataDec, !format(created_at, "%Y-%m-%d") == "2022-11-30")

#Remove irrelevant columns
toKeep <- c("created_at","tweet")
chatGPTDataDec <- chatGPTDataDec[names(chatGPTDataDec) %in% toKeep]

```

## b. 2023 data 

```{r}

#Convert time to date
chatGPTData2023$date <- as.POSIXct(chatGPTData2023$date, format = "%Y-%m-%d %H:%M:%S")

#Rename columns to match December dataset
chatGPTData2023 <- chatGPTData2023 %>% rename(created_at = date, tweet = content)

#Remove irrelevant columns
chatGPTData2023 <- chatGPTData2023[names(chatGPTData2023) %in% toKeep]

#Reorder to match December dataset
chatGPTData2023 <- chatGPTData2023 %>% select(created_at, tweet)

```

## c. March 2023 subset

```{r}

#Create smaller dataset selecting only March 2023 from second dataset 
chatGPTDataMar <- subset(chatGPTData2023, format(created_at, "%m") == "03")

```

## d. Dataset with all months combined

```{r}

#Merge dataset into one for time series plot
largerSet <- rbind(chatGPTData2023, chatGPTDataDec)

#Remove missing values from larger set
largerSet <- na.omit(largerSet)

```

# 3. Visualizations (larger dataset)

## a. Tweets by day

```{r}

#Number of tweets grouped by day
tweetsOverTimeDay <- largerSet %>% 
  group_by(day = date(created_at)) %>% count()

#Create the time series plot
ggplot(tweetsOverTimeDay, aes(x = day, y = n)) +
  geom_point(size = 0.5, color = "skyblue1") + 
  geom_line(color = "skyblue1") +
  geom_smooth(method = "lm", se = FALSE, color = "red2", size = 0.5) +
  theme_classic() +
  labs(title = "Tweets By Day", x = "", y = "Tweet Count") +
  scale_x_date(date_labels = "%b %d", date_breaks = "7 day") + 
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

#Top three days
topThree <- tail(order(tweetsOverTimeDay$n), 3)
topThreeValues <- tweetsOverTimeDay$n[topThree]
topThreeDays <- tweetsOverTimeDay$day[topThree]
topThreeDF <- data.frame(values = topThreeValues, day = topThreeDays)
print(topThreeDF)

```

## b. Tweets by month

```{r}

#Number of tweets grouped by month
tweetsOverTimeMonth <- largerSet %>% 
  group_by(month = month(created_at), year = year(created_at)) %>% count()

#Create a bar plot by month
ggplot(tweetsOverTimeMonth, aes(x = ymd(paste(year, month, "01", sep = "-")), y = n)) +
  geom_bar(stat = "identity", fill = "skyblue1") +
  geom_smooth(method = "lm", se = FALSE, color = "red2", size = 0.5) +
  theme_classic() +
  labs(title = "Tweets By Month", x = "Month/Year", y = "Tweet Count") +
  scale_x_date(date_labels = "%b %Y") +
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

# 4. Text wrangling

## a. Functions needed for text wrangling

```{r}

#Merges chat with gpt
removeChatGPT <- function(x)
{
  x <- gsub("(chat|gpt|chatgpt)", " ", x, ignore.case = TRUE)
}

#Remove twitter handles
removeTwitterHandles <- function(x) 
{
  gsub("@\\w+", " ", x)
}

#Remove any hyperlinks
removeHyperlinks <- function(x)
{
  gsub("https?://.+", " ", x)
}

#Remove punctuation
replacePunctuation <- function(x) 
{
  gsub("[[:punct:]]", " ", x)
}

replaceApostrophe <- function(x) 
{
  gsub("’", "", x)
}

tmMapCleanup <- function(x)
{
  
  #Convert text to all lowercase 
  print(as.character(x[[2]]))
  x <- tm_map(x, content_transformer(tolower))
  
  #Merge words chat and gpt
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(removeChatGPT))
  
  #Remove twitter handles
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(removeTwitterHandles))
  
  #Remove hyperlinks
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(removeHyperlinks))
  
  #Replace apostrophe
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(replaceApostrophe))
  
  #Replace punctuation
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(replacePunctuation))
  
  #Remove word chatGPT
  print(as.character(x[[2]]))
  x <- tm_map(x,content_transformer(removeChatGPT))
  
  #Remove stop words 
  print(as.character(x[[2]]))
  x <- tm_map(x, removeWords, stopwords())
  
  #Remove numbers  
  print(as.character(x[[2]]))
  x <- tm_map(x, removeNumbers)
  
  #Remove additional white space 
  print(as.character(x[[2]]))
  x <- tm_map(x, stripWhitespace)
  
  #Reduce words to root words by stemming 
  print(as.character(x[[2]]))
  x <- tm_map(x, stemDocument)
  
  print(as.character(x[[2]]))
  
  return(x)
}

```

## b. Apply text wrangling functions to data

```{r}

#Convert datasets to text corpus for text wrangling
tweetCorpusDec <- VCorpus(VectorSource(chatGPTDataDec$tweet))
tweetCorpusMar <- VCorpus(VectorSource(chatGPTDataMar$tweet))

#Apply clean up function
tweetCorpusMarClean <- tmMapCleanup(tweetCorpusMar)
tweetCorpusDecClean <- tmMapCleanup(tweetCorpusDec)

```

## c. Convert to term document matrix for word cloud and top ten visualizations

### i. December

```{r}

#Create term document matrix
fullTDMDec <- TermDocumentMatrix(tweetCorpusDecClean)

#Verify number of rows and columns, check first few terms
fullTDMDec$ncol
fullTDMDec$nrow
fullTDMDec$dimnames$Terms[1:3]

#Remove sparse terms to make dataset smaller (needed in order to convert into matrix)
smallTDMDec <- removeSparseTerms(fullTDMDec, sparse = 0.99)

#Verify number of rows and columns, check first few terms
smallTDMDec$ncol
smallTDMDec$nrow
smallTDMDec$dimnames$Terms[1:3]

```

### ii. March

```{r}

#Create term document matrix
fullTDMMar <- TermDocumentMatrix(tweetCorpusMarClean)

#Verify number of rows and columns, check first few terms
fullTDMMar$ncol
fullTDMMar$nrow
fullTDMMar$dimnames$Terms[1:3]

#Remove sparse terms to make dataset smaller (needed in order to convert into matrix)
smallTDMMar <- removeSparseTerms(fullTDMMar, sparse = 0.99)

#Verify number of rows and columns, check first few terms
smallTDMMar$ncol
smallTDMMar$nrow
smallTDMMar$dimnames$Terms[1:3]

```

# 4. Visualizations

## a. Word clouds 

### i. December

```{r}

#Convert small TDM to matrix
tdmMatrixDec <- as.matrix(smallTDMDec)

#Calculate row sums
wDec <- rowSums(tdmMatrixDec)

#Word cloud plot
set.seed(222)
wordcloud(words = names(wDec),
          freq = wDec,
          random.order = F,
          colors = brewer.pal(6, "Blues"),
          rot.per = 0.7,
          scale = c(3,0.8))
title(main = "December 2022")


```

### ii. March

```{r}

#Convert small TDM to matrix
tdmMatrixMar <- as.matrix(smallTDMMar)

#Calculate row sums
wMar <- rowSums(tdmMatrixMar)

#Word cloud plot
set.seed(222)
wordcloud(words = names(wMar),
          freq = wMar,
          random.order = F,
          colors = brewer.pal(6, "Blues"),
          rot.per = 0.7,
          scale = c(3,0.8))
title(main = "March 2022")

```

## b. Top ten words 

### i. December

```{r}

#Create data frame from words 
wordsDFDec <- data.frame(words = names(wDec), counts = wDec, row.names = 1:length(wDec))

#Arrange data frame in order and get top ten
wordsDFDecTop <- wordsDFDec %>% arrange(desc(counts)) %>% slice(1:10)

#Bar plot showing top ten words 
ggplot(data = wordsDFDecTop, aes(x = reorder(words, -counts), y = counts)) +
  geom_bar(stat = "identity", fill = "skyblue1") +
  theme_classic() +
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(size = 12), legend.position="bottom", legend.title = element_blank(), axis.text.y = element_text(size = 12)) +
  labs(title = "Top Ten Words - December 2022", x = "", y = "Count of Tweets")

```

### ii. March

```{r}

#Create data frame from words 
wordsDFMar <- data.frame(words = names(wMar), counts = wMar, row.names = 1:length(wMar))

#Arrange data frame in order and get top ten
wordsDFMarTop <- wordsDFMar %>% arrange(desc(counts)) %>% slice(1:10)

#Bar plot showing top ten words 
ggplot(data = wordsDFMarTop, aes(x = reorder(words, -counts), y = counts)) +
  geom_bar(stat = "identity", fill = "skyblue1") +
  theme_classic() +
  scale_y_continuous(labels = comma) +
  theme(axis.text.x = element_text(size = 12), legend.position="bottom", legend.title = element_blank(), axis.text.y = element_text(size = 12)) +
  labs(title = "Top Ten Words - March 2023", x = "", y = "Count of Tweets")

```


# 5. Sentiment analysis

## a. Calculate sentiment

```{r}

#Get sentiment for December
corpusTextDec <- sapply(tweetCorpusDecClean, content)
textDec <- as.character(corpusTextDec)
textDec[1:10]
textDec <- iconv(textDec)
sentimentScoreDec <- get_nrc_sentiment(textDec)

#Get sentiment for March
corpusTextMar <- sapply(tweetCorpusMarClean, content)
textMar <- as.character(corpusTextMar)
textMar[1:10]
textMar <- iconv(textMar)
sentimentScoreMar <- get_nrc_sentiment(textMar)

```

## b. Sentiment visualizations

### i. Plot emotions

```{r}

# Calculate column sums for both months
columnSumsDec <- data.frame(sort(colSums(prop.table(sentimentScoreDec[, 1:8]))))
columnSumsMar <- data.frame(sort(colSums(prop.table(sentimentScoreMar[, 1:8]))))

#Rename columns 
colnames(columnSumsDec) <- c("percentDec")
colnames(columnSumsMar) <- c("percentMar")

#Combine into one data frame for plotting
combinedColumnSums <- cbind(columnSumsDec, columnSumsMar)
combinedColumnSums$emotion <- row.names(combinedColumnSums)

#Plot emotions
meltCombinedColumnSums <- melt(combinedColumnSums)
ggplot(meltCombinedColumnSums, aes(x = emotion, y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +
    theme_classic() +
    labs(x = 'Emotion', y = 'Percent Tweets', title = "% Tweets by Emotion") +
    scale_y_continuous(labels = percent_format()) +
    scale_fill_manual(values = c("azure3", "skyblue1"), labels = c("December 2022", "March 2023")) + 
    theme(axis.text.x = element_text(size = 12), legend.position="bottom", legend.title = element_blank(), axis.text.y = element_text(size = 12))

```

### ii. Plot sentiment

```{r}
# Calculate column sums
columnSumsDecS <- data.frame(sort(colSums(prop.table(sentimentScoreDec[, 9:10]))))
columnSumsMarS <- data.frame(sort(colSums(prop.table(sentimentScoreMar[, 9:10]))))

#Rename columns
colnames(columnSumsDecS) <- c("percentDecS")
colnames(columnSumsMarS) <- c("percentMarS")

#Combined into one data frame for plotting
combinedColumnSumsS <- cbind(columnSumsDecS, columnSumsMarS)
combinedColumnSumsS$emotion <- row.names(combinedColumnSumsS)

#Plot emotion
meltCombinedColumnSumsS <- melt(combinedColumnSumsS)
ggplot(meltCombinedColumnSumsS, aes(x = emotion, y = value)) + 
    geom_bar(aes(fill = variable),stat = "identity",position = "dodge") +
    theme_classic() +
    labs(x = 'Emotion', y = 'Percent Tweets', title = "% Tweets by Sentiment") +
    scale_y_continuous(labels = percent_format()) +
    scale_fill_manual(values = c("azure3", "skyblue1"), labels = c("December 2022", "March 2023")) + 
    theme(axis.text.x = element_text(size = 12), legend.position="bottom", legend.title = element_blank(), axis.text.y = element_text(size = 12))

```

